Simulating the Lipkin-Meshov-Glikov model on a classical computer

### 2023-05-31 doc/paper.tex
Added subsections on one- and two-body Hamiltonians and mentioned the von neumann entropy for the two-body Hamiltonian in the theory part of the paper
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

Here we set the parameters $E_1 = 0$, $E_2 = 4$, $V_{11} = -V_{22} = 3$, and $V_{12} = V_{21} = 0.2$. The non-interacting solutions represent our computational basis. Pertinent to our choice of parameters, at $\lambda \geq \frac{2}{3}$, the lowest eigenstate is dominated by $|1\rangle$ while the upper is $|0\rangle$. At $\lambda = 1$, the $|0\rangle$ mixing of the lowest eigenvalue is 1\%, while for $\lambda \leq \frac{2}{3}$, we have a $|0\rangle$ component of more than 90\%. The character of the eigenvectors has therefore been interchanged when passing $\lambda = \frac{2}{3}$. The value of the parameter $V_{12}$ represents the strength of the coupling between the two states.

\subsubsection{Two-body Hamiltonian}

>>>>

Here we set the parameters $E_1 = 0$, $E_2 = 4$, $V_{11} = -V_{22} = 3$, and $V_{12} = V_{21} = 0.2$. The non-interacting solutions represent our computational basis. Pertinent to our choice of parameters, at $\lambda \geq \frac{2}{3}$, the lowest eigenstate is dominated by $|1\rangle$ while the upper is $|0\rangle$. At $\lambda = 1$, the $|0\rangle$ mixing of the lowest eigenvalue is 1\%, while for $\lambda \leq \frac{2}{3}$, we have a $|0\rangle$ component of more than 90\%. The character of the eigenvectors has therefore been interchanged when passing $\lambda = \frac{2}{3}$. The value of the parameter $V_{12}$ represents the strength of the coupling between the two states. ([Fys5419, project 1])

\subsubsection{Two-body Hamiltonian}
We also introduce a simple system composed of two qubits, where the system can be thought of as composed of two subsystems A and B.
Each subsystem has computational basis states
\begin{equation}
|0\rangle_{A,B} = \begin{bmatrix} 1 \\ 0 \end{bmatrix},
\end{equation}
\begin{equation}
|1\rangle_{A,B} = \begin{bmatrix} 0 \\ 1 \end{bmatrix}.
\end{equation}
The subsystems could represent single particles or composite many-particle
systems of a given symmetry. This leads to the many-body computational basis
states
\begin{equation}
|00\rangle = |0\rangle_A \otimes |0\rangle_B = \begin{bmatrix} 1 \\ 0 \\ 0 \\ 0 \end{bmatrix},
\end{equation}
\begin{equation}
|01\rangle = |0\rangle_A \otimes |1\rangle_B = \begin{bmatrix} 0 \\ 1 \\ 0 \\ 0 \end{bmatrix},
\end{equation}
\begin{equation}
|10\rangle = |1\rangle_A \otimes |0\rangle_B = \begin{bmatrix} 0 \\ 0 \\ 1 \\ 0 \end{bmatrix},
\end{equation}
and finally
\begin{equation}
|11\rangle = |1\rangle_A \otimes |1\rangle_B = \begin{bmatrix} 0 \\ 0 \\ 0 \\ 1 \end{bmatrix}.
\end{equation}
These computational basis states also define the eigenstates of the non-interacting Hamiltonian
\begin{equation}
H_0 |00\rangle = \epsilon_{00} |00\rangle,
\end{equation}
\begin{equation}
H_0 |10\rangle = \epsilon_{10} |10\rangle,
\end{equation}
\begin{equation}
H_0 |01\rangle = \epsilon_{01} |01\rangle,
\end{equation}
and
\begin{equation}
H_0 |11\rangle = \epsilon_{11} |11\rangle.
\end{equation}
The interacting part of the Hamiltonian \(H_I\) is given by the tensor product of
two \(\sigma_x\) and \(\sigma_z\) matrices, respectively, that is
\begin{equation}
H_I = H_x \sigma_x \otimes \sigma_x + H_z \sigma_z \otimes \sigma_z,
\end{equation}
where \(H_x\) and \(H_z\) are interaction strength parameters. Our final Hamiltonian
matrix is given by
\begin{equation}
H = \begin{bmatrix}
\epsilon_{00} + H_z & 0 & 0 & H_x \\
0 & \epsilon_{10} - H_z & H_x & 0 \\
0 & H_x & \epsilon_{01} - H_z & 0 \\
H_x & 0 & 0 & \epsilon_{11} + H_z
\end{bmatrix}.
\end{equation}

\subsubsection{Von Neumann entropy for the two-body Hamiltonian}
The four eigenstates of the above Hamiltonian matrix can in turn be used to define density matrices. As an example, the density matrix of the first eigenstate (lowest energy \(E_0\), denoted as \(\Psi_0\)) is

\begin{equation}
\rho_0 = (\alpha_{00}|00\rangle\langle00| + \alpha_{10}|10\rangle\langle10| + \alpha_{01}|01\rangle\langle01| + \alpha_{11}|11\rangle\langle11|)
\end{equation}

where the coefficients \(\alpha_{ij}\) are the eigenvector coefficients resulting from the solution of the above eigenvalue problem.

We can then in turn define the density matrix for the subsets A or B as

\begin{equation}
\rho_A = \text{Tr}_B(\rho_0) = \langle0|\rho_0|0\rangle_B + \langle1|\rho_0|1\rangle_B,
\end{equation}

or

\begin{equation}
\rho_B = \text{Tr}_A(\rho_0) = \langle0|\rho_0|0\rangle_A + \langle1|\rho_0|1\rangle_A.
\end{equation}

The density matrices for these subsets can be used to compute the so-called von Neumann entropy, which is one of the possible measures of entanglement. A pure state has entropy equal to zero, while entangled states have an entropy larger than zero. The von Neumann entropy is defined as

\begin{equation}
S(A, B) = -\text{Tr}(\rho_{A,B} \log_2(\rho_{A,B})).
\end{equation}

We will use this definition, to measure the level of entanglement present in the system as a function of the interatction strength parameter $\lambda$.
